DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-10-04 paper/evaluation.tex
eval. E. bandwidth revision

In each experiment, we run a single DNN model one core, and
co-scheduled an increasing number of the memory intensive synthetic
benchmark on the remaining idle cores.
Figure~\ref{fig:}
The effects of the benchmarks can seen in Figure 9 and Table VI.

\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth]{figs/BenchmarkChart}
  \caption{Effect of benchmarks on inferencing time.}
  \label{fig:}
\end{figure}

\begin{table}[h]
\centering
  \begin{tabular} {| l | l | l | l | l | l |}
  \hline
  \textbf{num cores} & 1 & 2 & 3 \\ \hline
  \textbf{L1 refs} & 3.16E+10 & 3.17E+10 & 3.06E+10 \\ \hline
  \textbf{L1 misses} & 5.41E+08 & 5.35E+08 & 4.98E+08 \\ \hline
  \textbf{L1 miss \%} & 1.71 & 1.69 & 1.62 \\ \hline
  \textbf{L2 refs} & 3.49E+09 & 3.23E+09 & 3.21E+09 \\ \hline
  \textbf{L2 misses} & 8.05E+08 & 7.54E+08 & 4.98E+08 \\ \hline
  \textbf{L2 miss \%} & 23.05 & 23.37 & 15.51 \\ 
  \hline
  \end{tabular}
  \caption{ Effect of benchmarks on cache accesses. }
\end{table}

The presence of the benchmark(s) had a noticable effect on the DeepPicar, with all tests showing 
increases in inferencing operations. The least change happened when the model was run on 3 cores, and a 
single benchmark was run on the final core. Even with a slight increase in the number of L2 cache 
misses, inferencing was able to be completed in under 100 ms. The other tests, however, showed dramatic 
changes when the model was run on 1 and 2 cores. In both instances, L2 misses rose to $\sim$23\% of all 
references. The additional benchmark in the 1 core test was even more detrimental as the average 
inferencing time was at least 150 ms longer than when the model used 2 cores.

With the introduction of synthetic benchmarks, the DeepPicar failed to comply with the 50 ms WCET across 
all tests run. Even in the best case, inferencing operations take twice as long as the deadline to 
complete. As a result, if benchmarks were required to run during autonomous operation, the DeepPicar 
wouldn't be capable of meeting its deadlines.

\subsection{Summary}
We found that the DeepPicar is capable of successfully completing all necessary inferencing 
operations within a given deadline of 50 ms. Between these operations, it was discovered that the 
angle prediction time of the model was the dominating step in the processing time of a single frame. 
The other operations were found to execute in very little time, with either one taking 5 ms at most.

When running a single model, DeepPicar performed the best when it used all 4 cpu cores, as it was 
able to process a frame in 43 ms on average. This feat can still be accomplished when running a 
single model on 3 cpu cores, and is also possible when using 2 cores. The only time in which the Pi 
can't run a single model under 50 ms is when only 1 core is used. This was also the case when 
multiple models were run concurrently as no models were able to consistently complete inference 
operations in under 50 ms.

However, the DeepPicar was shown to be capable of handling other potential WCETs. If given a larger 
deadline, the capabilities of DeepPicar as an autonomous vehicle platform would be greatly increased. 
For example, if tested with a deadline of 66.67 ms, or 15 fps, the platform would have passed several 
of the experiments performed. In actuality, the DeepPicar would have only been unsatisfactory in the 
4 model test where it would have missed the deadlines by $\sim$10 ms, and the benchmark tests, where 
it would still miss by $\sim$33 ms at best.

@@

The most notable issue that can affect the cpu clock speed is that of the power supplied to the 
Raspberry Pi. In essence, it is necessary that the Pi be supplied with 2 Amps, as any less could 
hinder the Pi's ability to maintain a 1.2 GHz frequency. In experiments conducted with a power supply that 
only provided 1 Amp, the Pi was unable to sustain a 1.2 GHz clock speed and, instead, fluctuated 
between operating at 600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly 
recommended, that the power supply used for the Raspberry Pi 3 be capable of outputting 2 Amps, 
otherwise optimal performance isn't guaranteed.

Another factor that can affect clock speed is that of the cpu's temperature. Some model operations can 
be computationally intensive, thus it is possible for the temperature of the cpu to become relatively 
high. This can be especially problematic in situations where multiple models are running 
simultaneously on the Pi. Consequently, thermal throttling may be used to decrease the clock 
speed so that the cpu temperature stays at a safe level. Thus the Raspberry Pi may not be suited 
for prolonged use, especially in cases where the workload is relatively larger, such as running multiple 
models. Rather, the Pi seems to be better suited for running in set periods, after which it is turned 
off or made idle so that the cpu is allowed time to cool down.

>>>>

In each experiment, we run a single DNN model one core, and
co-scheduled an increasing number of the memory intensive synthetic
benchmarks~\footnote{We use the \emph{Bandwidth} benchmark in the
  IsolBench suite, with the following command line parameters: \texttt{\$
  bandwdith -a write -m 16384}}, on the remaining idle cores.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth}
  \caption{Effect of memory performance hogs on the DNN
    inferencing. The DNN model uses Core 0 and memory-hog co-runnes
    use the rest of the cores.}
  \label{fig:}
\end{figure}

%% \begin{table}[h]
%% \centering
%%   \begin{tabular} {| l | l | l | l | l | l |}
%%   \hline
%%   \textbf{num cores} & 1 & 2 & 3 \\ \hline
%%   \textbf{L1 refs} & 3.16E+10 & 3.17E+10 & 3.06E+10 \\ \hline
%%   \textbf{L1 misses} & 5.41E+08 & 5.35E+08 & 4.98E+08 \\ \hline
%%   \textbf{L1 miss \%} & 1.71 & 1.69 & 1.62 \\ \hline
%%   \textbf{L2 refs} & 3.49E+09 & 3.23E+09 & 3.21E+09 \\ \hline
%%   \textbf{L2 misses} & 8.05E+08 & 7.54E+08 & 4.98E+08 \\ \hline
%%   \textbf{L2 miss \%} & 23.05 & 23.37 & 15.51 \\ 
%%   \hline
%%   \end{tabular}
%%   \caption{ Effect of benchmarks on cache accesses. }
%% \end{table}

Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
of the DNN model running on the Core 0 as a function of the number of
co-scheduled memory intensive synthetic benchmarks. First, as we
increase the number of co-runners, the DNN model's execution times are
increased---by up to 9.6X---even though the DNN model is running on a
dedicated core (Core0). On the other hand, the DNN model's L2
cache-miss rates are not increase as much. This suggests that
the DNN model's exeuction increase cannot be fully explained by
increased L2 cache-misses. Instead, as we hypothesized in the previous
experiment, the increased memory pressure from the co-scheduled memory
intensive benchmarks is the primary cause of the DNN model's execution
time increase. Therefore, we conclude that DeepPicar's DNN model is
more senstive to DRAM access latency than L2 cache space.

This observation suggests that shared cache partitioning
techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
solutions for DeepPicar as its AI workload is more sensitive to memory
performance. Instead, memory controller focused isolation solutions,
either hardware or software-based ones (e.g.,~\cite{Guo2017,Yun2013}),
may be more important. Although our observation is made on a single
hardware platform running on a single DNN workload, we suspect that
many AI workloads may exhibit similar characteristics.

%% The presence of the benchmark(s) had a noticable effect on the
%% DeepPicar, with all tests showing increases in inferencing
%% operations. The least change happened when the model was run on 3
%% cores, and a single benchmark was run on the final core. Even with a
%% slight increase in the number of L2 cache misses, inferencing was able
%% to be completed in under 100 ms. The other tests, however, showed
%% dramatic changes when the model was run on 1 and 2 cores. In both
%% instances, L2 misses rose to $\sim$23\% of all references. The
%% additional benchmark in the 1 core test was even more detrimental as
%% the average inferencing time was at least 150 ms longer than when the
%% model used 2 cores.

%% With the introduction of synthetic benchmarks, the DeepPicar failed to
%% comply with the 50 ms WCET across all tests run. Even in the best
%% case, inferencing operations take twice as long as the deadline to
%% complete. As a result, if benchmarks were required to run during
%% autonomous operation, the DeepPicar wouldn't be capable of meeting its
%% deadlines.

\subsection{Summary}
We found that the DeepPicar is capable of successfully completing all
necessary inferencing operations within a given deadline of 50
ms. Between these operations, it was discovered that the angle
prediction time of the model was the dominating step in the processing
time of a single frame. The other operations were found to execute in
very little time, with either one taking 5 ms at most. When running a
single model, DeepPicar performed the best when it used all 4 cpu
cores, as it was able to process a frame in 43 ms on average. This
feat can still be accomplished when running a single model on 3 cpu
cores, and is also possible when using 2 cores. The only time in which
the Pi can't run a single model under 50 ms is when only 1 core is
used. This was also the case when multiple models were run
concurrently as no models were able to consistently complete inference
operations in under 50 ms. 
However, the DeepPicar was shown to be capable of handling other
potential WCETs. If given a larger deadline, the capabilities of
DeepPicar as an autonomous vehicle platform would be greatly
increased. For example, if tested with a deadline of 66.67 ms, or 15
fps, the platform would have passed several of the experiments
performed. In actuality, the DeepPicar would have only been
unsatisfactory in the 4 model test where it would have missed the
deadlines by $\sim$10 ms, and the benchmark tests, where  
it would still miss by $\sim$33 ms at best.

@@

The most notable issue that can affect the cpu clock speed is that of
the power supplied to the Raspberry Pi. In essence, it is necessary
that the Pi be supplied with 2 Amps, as any less could hinder the Pi's
ability to maintain a 1.2 GHz frequency. In experiments conducted with
a power supply that only provided 1 Amp, the Pi was unable to sustain
a 1.2 GHz clock speed and, instead, fluctuated between operating at
600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly
recommended, that the power supply used for the Raspberry Pi 3 be
capable of outputting 2 Amps, otherwise optimal performance isn't
guaranteed.  Another factor that can affect clock speed is that of the
cpu's temperature. Some model operations can be computationally
intensive, thus it is possible for the temperature of the cpu to
become relatively high. This can be especially problematic in
situations where multiple models are running  simultaneously on the
Pi. Consequently, thermal throttling may be used to decrease the clock
speed so that the cpu temperature stays at a safe level. Thus the
Raspberry Pi may not be suited for prolonged use, especially in cases
where the workload is relatively larger, such as running multiple
models. Rather, the Pi seems to be better suited for running in set
periods, after which it is turned  off or made idle so that the cpu is
allowed time to cool down.
